Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-07-10 paper/tci.tex
Changes in Fig.1, related work, Intro.

However, in the last decade, reconstruction from reduced views has
been made possible by assuming the data to exhibit sparsity under
certain mathematical transforms $\boldsymbol{\Upsilon}$ such as the
wavelet transforms, or the Discrete Cosine Transform (DCT). This is
known as sparsity prior and is the fundamental principle in the widely
used Compressed Sensing (CS) technique~\cite{Donoho,introCS}.
There are multiple ways to incorporate the sparsity prior using CS. We use the LASSO (least absolute shrinkage and selection operator) which iteratively solves for the solution by penalizing a combination of least squares error and $L_1$-norm of the sparse coefficients $\boldsymbol{\theta}$ of the object $\boldsymbol{x}$. If $\boldsymbol{\mathcal{R}}$ represents the acquisition model, then the LASSO solution is described by one that minimizes
 \begin{equation}
J_{\text{CS}}(\boldsymbol{\theta}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1
\label{Eq:simple_cs}
 \end{equation}
 We solve this cost function using the popular l1-regularized least
 squares (l1-ls) package available in~\cite{l1ls}, and choose DCT as
 our sparsity basis $\boldsymbol{\Upsilon}$.  Fig.~\ref{fig:story}
 (second column) demonstrates the benefit of using sparsity prior when
 the number of projection views is limited (100 views).

@@

 \section{Related Work}
\label{sec:related}
Few view projection has been explored in~\cite{Essam2015,Polak2017}
and~\cite{Li2015}. However
{\color{red} add here!!  Why is this bad and how does it compare with
ours? }

Most object-prior based reconstruction has been in the areas of
dynamic CT and 4D-CT. In dynamic CT, the object being scanned
undergoes changes between successive projection views and hence is not
stationary even within a single complete scan. In contrast, in 4D-CT,
a set of 3D CT scans over short intervals of time are taken as the
object undergoes changes between successive scans. In both these
cases, it may be possible to use very specific object-properties for
reconstruction. For example, in~\cite{Van2015}, the knowledge of the
attenuation coefficient of the fluid is used as a prior for
reconstructing its flow through gravel. In~\cite{Koen2020}, the
changes across the sucessive scan volumes is assumed to be continuous,
thus enabling the use of optical flow to model the motion between
corresponding voxels of different scan volumes. In another
study~\cite{vincent2017}, changes between successive scans were
modelled by affine deformation whose parameters were computed for
motion-correction. Such a correction enables acquisition from fewer
views than would have otherwise not been possible for reconstruction
with adequate fidelity. In~\cite{daniil2015}, all the scan volumes are
reconstructed together using spatio-temporal regularization.

@@

\begin{table}[!h]
  \centering
\caption{Computation time for reconstruction of volumes using various methods. The configuration of the system used was x86\_64 architecture, 62GB RAM, 12-core AMD 3.8GHz processor, with Nvidia GeForce RTX 2080 Ti GPU.}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
   & \textbf{FDK} & \textbf{l1-ls} & \textbf{\begin{tabular}[c]{@{}c@{}}Uniform\\prior + l1-ls\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Spatially-Varying\\prior + l1-ls\end{tabular}} \\\hline
%\textbf{\begin{tabular}[c]{@{}c@{}}Potato\\size:(150,150,100)\\45 views\end{tabular}}  & 5 sec. & 8 min.  & 54 min. & 53 min. \\ \hline
\textbf{\begin{tabular}[c]{@{}c@{}}Okra\\size:(338,338,123)\\45 views\end{tabular}} & 6 sec. & 8 min.  & 25 min. & 27 min. \\ \hline
\textbf{\begin{tabular}[c]{@{}c@{}}Sprouts\\size:(130,130,130)\\xx views\end{tabular}}  &  &   &  &  \\ \hline
\end{tabular}
\label{table:Computation_time_3D}
\end{table}

>>>>

However, in the last decade, reconstruction from reduced views has
been made possible by assuming the data to exhibit sparsity under
certain mathematical transforms $\boldsymbol{\Upsilon}$ such as the
wavelet transforms, or the Discrete Cosine Transform (DCT). This is
known as sparsity prior and is the fundamental principle in the widely
used Compressed Sensing (CS) technique~\cite{Donoho,introCS}.  There
are multiple ways to incorporate the sparsity prior using CS. We use
the LASSO (least absolute shrinkage and selection operator) which
iteratively solves for the solution by penalizing a combination of
least squares error and $L_1$-norm of the sparse coefficients
$\boldsymbol{\theta}$ of the object $\boldsymbol{x}$. If $\boldsymbol{\Psi}$
represents the sparsity basis i.e. if $\boldsymbol{x} = \boldsymbol{\Psi\theta}$, and 
$\boldsymbol{\mathcal{R}}$ represents the acquisition model, then the
LASSO solution is described by one that minimizes
 \begin{equation}
J_{\text{CS}}(\boldsymbol{\theta}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1
\label{Eq:simple_cs}
 \end{equation}
 We solve this cost function using the popular l1-regularized least
 squares (l1-ls) package available in~\cite{l1ls}, and choose DCT as
 our sparsity basis $\boldsymbol{\Upsilon}$.  Fig.~\ref{fig:story}
 (second column) demonstrates the benefit of using sparsity prior when
 the number of projection views is limited (100 views).

@@

 \section{Related Work}
\label{sec:related}
There exist techniques that aim to reconstruct from few-views in
specialized applications. For example, in~\cite{Essam2015},
sparsity-constrained optimization is presented for angiography. Here,
the regions of interest are the vessels alone and they are highlighted
by physically inserting a contrast agent, and therefore there is a
inherent sharper contrast between the vessels and the
background. In many other general application scenarios like the ones presented in this paper,
any information about the regions of interest is not known. In other applications where the spatial-gradient of the underlying volume is known to be sparse, the `total variation' method, as used in~\cite{Li2015,Polak2017} can be applied. As we will later see in Figs.\ref{fig:object-prior_test_okra} and~\ref{fig:object-prior_test_sprouts}, the spatial-gradient of many datasets is not sparse.

Most object-prior based reconstruction has been in the areas of
dynamic CT and 4D-CT. In dynamic CT, the object being scanned
undergoes changes between successive projection views and hence is not
stationary even within a single complete scan. In contrast, in 4D-CT,
a set of 3D CT scans are taken over short intervals of time and all the 3D volumes are reconstructed together. Here, the object undergoes changes between successive scans. ~\cite{PICCS} is one of the earliest work to establish the use of object-prior along with sparsity prior in dynamic CT. Its limitation was the unchecked over-emphasis of the prior on the reconstruction of the test at hand. This limitation was later dealt with by applying very specific object-properties for
reconstruction. For example, in~\cite{Van2015}, the knowledge of the
attenuation coefficient of the fluid is used as a prior for
reconstructing its flow through gravel. In~\cite{Koen2020}, the
changes across the sucessive scan volumes is assumed to be continuous,
thus enabling the use of optical flow to model the motion between
corresponding voxels of different scan volumes. In another
study~\cite{vincent2017}, changes between successive scans were
modelled by affine deformation whose parameters were computed for
motion-correction. Such a correction enables acquisition from fewer
views than would have otherwise not been possible for reconstruction
with adequate fidelity. In~\cite{daniil2015}, all the scan volumes are
reconstructed together using spatio-temporal regularization.

@@

\begin{table}[!h]
  \centering
\caption{Computation time for reconstruction of volumes using various methods. The configuration of the system used was x86\_64 architecture, 62GB RAM, 12-core AMD 3.8GHz processor, with Nvidia GeForce RTX 2080 Ti GPU.}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
   & \textbf{FDK} & \textbf{l1-ls} & \textbf{\begin{tabular}[c]{@{}c@{}}Uniform\\prior + l1-ls\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Spatially-Varying\\prior + l1-ls\end{tabular}} \\\hline
%\textbf{\begin{tabular}[c]{@{}c@{}}Potato\\size:(150,150,100)\\45 views\end{tabular}}  & 5 sec. & 8 min.  & 54 min. & 53 min. \\ \hline
\textbf{\begin{tabular}[c]{@{}c@{}}Okra\\size:(338,338,123)\\45 views\end{tabular}} & 6 sec. & 8 min.  & 25 min. & 60 min. \\ \hline
\textbf{\begin{tabular}[c]{@{}c@{}}Sprouts\\size:(130,130,130)\\xx views\end{tabular}}  &  &   &  &  \\ \hline
\end{tabular}
\label{table:Computation_time_3D}
\end{table}
